---
title: "Assignment 1"
author: "YOUR NAME HERE"
date: "due 9/8/21"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Assignment 1
This is due prior to the beginning of live session on the due date. Please submit two files to the LMS: 1. a knit file (I prefer pdf, but will accept html - just choose one) and 2. your .Rmd code file. Round all reported statistics to the nearest hundredths place (i.e., two decimal places) when writing up conclusions/interpretations (no need to round them in code). Make sure you answer *all* parts of each question. Also, be sure to review the final knit document - if there are typos/formatting issues, please correct those before uploading. 

Please identify students with whom you worked on this assignment here (max of FOUR to a group):



## 1. [Week 1] Researchers at the University of Utah use a driving simulator in experiments on the effects of divided attention. In one of their experiments, participants were randomly assigned to one of two conditions. One condition was called the single-task condition, and participants’ task was to follow a course in the driving simulator. The other condition was called the dual-task condition, and participants’ had to follow the course in the driving simulator while at the same time engaged in a hands-free cell phone conversation. The driving simulator recorded each participant’s reaction time (in milliseconds) to an obstacle that appeared unexpectedly. The average reaction time was significantly greater in the dual-task condition. Therefore, divided attention had a detrimental effect on driving performance.

### a. Identify the independent variable in the study. Is this variable continuous or discrete? 



### b. Identify the dependent variable in the study. Is this variable continuous or discrete?



## 2. [Week 2] Consider the following frequency distribution:

Score | Frequency
------------- | -------------
7 | 18
6 | 77
5 | 20
4 | 7
3 | 3
2 | 1
1 | 3

### a. What is the sample size?



### b. What percent of participants had a score of 4?



### c. What is the percentile rank for a score of 5? 



### d. What score corresponds to the 80th percentile?



## 3. [Week 1] Define objects for values 5 and 1. *Use the objects* to compute the following:

### a. 5+1
```{r}

```

### b. 5^2
```{r}

```

### c. $\sqrt{5-1}$
```{r}

```

## 4. [Week 1] A math teacher asked 18 of her students to rate how much they like math on a 1 - 7 Likert-scale (1 = not at all, 7 = very much). She received the following responses: 3, 7, 5, 3, 5, 6, 3, 2, 6, 4, 4, 3, 4, 2, 3, 6, 4, 7.

### a. Enter the data into R as a vector with a name (of your choosing). Print the vector.
```{r}

```

### b. Create a histogram.
```{r}

```

### c. Describe the shape of the distribution using terminology learned in class.



## 5. [Week 1] The teacher also collected the final average in her math class for each of the 18 students: 84, 97, 88, 73, 81, 85, 63, 76, 85, 79, 83, 66, 76, 66, 82, 73, 81, 92

### a. Enter the data into R as a vector with a name. Print the vector.
```{r}

```

### b. Create a histogram.
```{r}

```

### c. Describe the shape of the distribution using terminology learned in class.



## 6. [Week 1] Create a data frame using the two vectors created in problems (4) and (5). Print the data frame.
```{r}

```


## 7. [Week 2] Import the undergrad student survey dataset (filename: survey.txt) into R. 

### a. Using R, calculate and report the mean, median, and standard deviation for each of the following variables: high school GPA (HSGPA) and college GPA (CollegeGPA).  
```{r}

```

### b. Interpret the meaning of the mean, median, and standard deviation of the high school GPA variable in the *context of this problem.* Do not provide general definitions here, but interpret what each statistic means for these data.



### c. Interpret the meaning of the mean, median, and standard deviation of the college GPA variable in the *context of this problem.* Do not provide general definitions here, but interpret what each statistic means for these data.



### d. Based on the descriptive statistics interpreted in parts (c) and (d), write a sentence or two about the difference between college and high school GPA for this sample.


