```{r}
REMOVE ME
```
---
title: "Week 11 - Correlation"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list=ls())

knitr::opts_chunk$set(echo = TRUE)
```

This week we will learn how to compute and analyze correlations

Let's look at the phd data from async. First, import the data.
```{r}
phd_data <- read.table("week11_phd.txt", header=TRUE, sep="\t")
```
LOOK at your data after you import it! How many vars? How many obs? Note variable names, etc.

Now let's make make a scatterplot (IV on x-axis; DV on y-axis)
```{r}
plot(phd_data$Time, phd_data$Pubs, 
     main="Scatterplot of Time & Publications",
     xlab="Time Since PhD Grad (in years)", 
     ylab="Publications", pch=12)
```

We can use R to compute the Pearson Correlation and obtain other values for the hypothesis test
```{r}
cor.test(phd_data$Time, phd_data$Pubs, method=c("pearson"))
```

Calculate effect size (R^2)
```{r}
0.6566546^2
```

Now, let's look the example from asyc for the spearman correlation.
```{r}
english_data <- read.table("week11_english.txt", header=TRUE, sep="\t")

cor.test(english_data$Words, english_data$Grade, method=c("spearman"))
```

Point-Biserial Correlation (note that this code is the same as Pearson's, but the data setup differs)
```{r}
byod_data <- read.table("week11_byod.txt", header=TRUE, sep="\t")

cor.test(byod_data$Byod, byod_data$Grade) #pt-biserial

#calculate effect size - R^2
0.378822^2
```

Phi Correlation (note that this code is the same as Pearson's, but the data setup differs)
```{r}
grad_data <- read.table("week11_graduate.txt", header=TRUE, sep="\t")

table(grad_data)

cor.test(grad_data$Business, grad_data$Doctorate) #phi 

#calculate effect size
(-0.2228072)^2
```

