```{r}
REMOVE ME 
```
---
title: "Week 6 - One-sample t-test"
author: "Lacey Hartigan"
date: "2/24/21"
output:
  pdf_document: default
---

```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#NOTE: if you don't have tidyverse installed yet, you need to install it. 
#Type the following code into your console and hit "enter": install.packages('tidyverse')
library(tidyverse)
```

## One-sample T-Test in R

Let's use R to perform calculations for the "couples" example, comparing the amount of time American couples spend discussing marital problems to the amount of time European couples spend discussing marital problems. 

First, let's write out our Null and Alternative hypotheses (in words AND in symbols):

\[
\begin{aligned}
H_0: \mu_{American} = \mu_{European} \\
\end{aligned}
\]
American couples spend the same amount of time discussing marital problems as European couples.
\[
\begin{aligned}
H_1: \mu_{American} \neq \mu_{European} \\ 
\end{aligned}
\]
#American couples spend a different amount of time discussing marital problems as compared to Euro couples.

First, let's calculate the t-statistic "by hand"
```{r}
(2.6-3)/(1.9/sqrt(70))
```
Next, we can calculate the p-value. Instead of `pnorm`, we will use `pt`.
The only difference is that  have to specify the df. We still have to multiply the probability for two-tailed tests (note: the t-table makes this easier). We still need to be careful with the lower.tail command (because we have a negative t-statistic we need to use TRUE).

```{r}
2*pt(-1.76,df=69, lower.tail=TRUE)

#or we could have used the positive value (because of symmetry)
2*pt(1.76,df=69, lower.tail=FALSE)
```
Decision: Fail to reject the H0.
Conclusion: A one-sample t-test revealed that there was no SS difference between American Couples (M=2.6) and Euro couples (M=3) on time spent discussing marital problems; t(69) = -1.76, p > 0.05.


Now, let's learn how to conduct a t-test when given a dataset. Let's look at sample data given for example 9.1 in the book. We're going to import this data (so taht we get the practice). 

This file is a .csv (comma separated values) file.
```{r}
#Import the file
OCD<-read_csv(file="OCD_data.csv")
OCD

#Run a one-sample t-test
t.test(OCD,mu=77.43, conf.level=.95)
```

Decision: Reject the null, b/c p<.05. So, BECAUSE we rejected the null, we can now use R to calculate the effect size "by hand."

```{r}
#Calculate Cohen's d:
(62-77.43)/20.94
```

FINDINGS/CONCLUSION (in APA style):
A one-sample two-tailed t-test at alpha level .05 revealed a statistically significant difference between our sample mean (M=62) and our population (Mu=77.43), t(17) = -3.13, p < 0.05, d = -0.74. Cohen's d signified a moderate effect. 

Set-up for reporting stats in APA (one-sample t-test): 
t(DF)=TSTAT, p(lessthanorgreaterthan).05, [if significant] d=COHENSVALUE.






### BONUS

If we wanted to get really fancy, we could write our own function to calculate cohen's d. This way, we can use this again and again, for one-sample t-tests. 
```{r}
Mu=77.43

cohensd <- function(x) {
  result <- (mean(x)-Mu)/sd(x)
}

d<-cohensd(OCD$OCD)
d
```



