---
title: "Week 8 - Confidence Intervals for T-Tests"
author: "Lacey Hartigan"
output:
  pdf_document: default
---
  
```{r include=FALSE}
rm(list=ls())

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

This week we show how to build confidence intervals in R for t-tests. There is actually no new code here because, by default, t.test() gives the CI in the output.

First, let's revisit the one-sample t-test example from last week, where we were given a population Mu of 77.43
```{r}
#load the data
OCDdata <- read.table("OCD.txt", header=TRUE, sep="\t")
```

```{r}
#run the one sample t-test
t.test(OCDdata$OCD, mu=77.43, conf.level=.95)
```
Q: Using the CI method - would we reject/fail to reject the null? Why?

A: We would reject the H0 because our population mu (77.43) was outside our confidence interval [51.59, 72.41]. NOTE how we display confidence intervals (using brackets and a comma in between the lower and upper bound).

Now, let's write the interpretation for the CI [51.59, 72.41].
Interpretation: We are 95% confident that the CI [51.59, 72.41] contains the true population mean for OCD. 
NOTE: We can interpret the CI whether or not we reject/fail to reject (interpretation of the CI doesn't affect rejection and vice versa.

Now, let's look at an Independent samples t-test -- two ways (based on data structure)
#1
```{r}
#load the data
consumed1<-read.table("caloriesnum.txt", header=T, sep="\t")
#run the test
t.test(consumed1$Slow, consumed1$Fast, var.equal=T)
```

#2
```{r}
#load the data
consumed2<-read.table("calorieschar.txt", header=T, sep="\t")
#run the test
t.test(consumed2$Calories ~ consumed2$Group, var.equal=T)
```

Q: Using the CI method - would we reject/fail to reject the null? Why?

A: We would fail to reject b/c the CI contains 0. 

Interpretation: We are 95% confident that the CI [-134.19, 234.19] contains the true mean difference. That is, if we took 100 samples and constructed CIs in the same way, 95 of them would contain the true mean difference.

APA conclusion: An independent samples t-test revealed no statistically significant difference between the fast (M=650, SD=130.38) and slow (M=600, SD=154.92) eaters on calorie consumption; t(10)=0.60, p>0.05.

Related samples t-test (process is the same; code differs slightly)
```{r}
grades<-read.table("grades.txt", header=T, sep="\t")

t.test(grades$Posttest, grades$Pretest, paired=T)
```

Q: Using the CI method - would we reject/fail to reject the null? Why?
A: We would reject the null, because the CI [2.11, 10.56] does not contain 0.

Interpretation: We are 95% confident that the true mean difference lies within the CI [2.11, 10.56].

Since we found significance, what would our next step be?
Calculate cohen's d
```{r}
#Dbar/SDdiff (recall how we found these last week)
6.33/5.5 #d = 1.15 large effect
```

```{r}
#pull descriptive stats for our write-up
mean(grades$Pretest)
sd(grades$Pretest)
mean(grades$Posttest)
sd(grades$Posttest)
```

APA conclusion: A related samples t-test revealed a statistically significant difference between pre-test (M=70.22, SD=9.22) and post-test (M=76.56, SD=9.30) grades; t(8)=3.45, p<0.05, d=1.15. Cohen's d signified a large effect.






