```{r}
REMOVE ME
```
---
title: "Week 12 - Linear Regression"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list=ls())

knitr::opts_chunk$set(echo = TRUE)
```

This week we will learn how to conduct a SLR (simple linear regression; aka ONE predictor variable) in R. 

Let's look at the phd data from async. First, import the data.
```{r}
phd_data <- read.table("week12_phd.txt", header=TRUE, sep="\t")
```
Remember - LOOK AT THE DATA after you read it in!!!

First, let's make a scatterplot to examine the data visually.
```{r}
plot(phd_data$Time, phd_data$Pubs ,
     main="Scatterplot of Time & Pubs", 
     xlab="Time Since Grad (in Years)", ylab="Publications", pch=9)
```

Let's run an Simple Linear Regression (SLR) with Pubs as DV and Time as IV; we will save it to an object called phd.regression
```{r}
#REGRESSIONOBJECTNAME <- lm(YVARIABLE ~ XVARIABLE, data=DATASETNAME)
phd.regression <- lm(Pubs ~ Time, data=phd_data)
```

Now pull the results from R's memory:
```{r}
#Parameter estimates
phd.regression

#More information for hypothesis tests
summary(phd.regression)

#confidence intervals for parameter estimates
confint(phd.regression)
```

What if we wanted to plot the regression line on our scatterplot? R makes that easy. 
```{r}
plot(phd_data$Time, phd_data$Pubs ,
     main="Scatterplot of Time & Pubs", 
     xlab="Time Since Grad (in Years)", ylab="Publications", pch=9)
abline(phd.regression, col="blue")
```

Now let's do another example (using data that is stored in R). You should ALWAYS look at the data first. Let's look at the mtcars dataset that comes with R.
```{r}
summary(mtcars)
```

Let's examine a possible relationship between car weight & miles per gallon
```{r}
plot(mtcars$wt, mtcars$mpg, 
     main="Scatterplot of Weight and MPG", 
     xlab="Weight in 1000s", ylab="MPG", pch=19)
```
Does it look like there may be a relationship? Why or why not?

Now let's see if we can predict MPG from car weight.
```{r}
cars.regression <- lm(mpg ~ wt, data=mtcars)
summary(cars.regression)
```

Add the regression line of best fit.
```{r}
plot(mtcars$wt, mtcars$mpg, 
     main="Scatterplot of Weight and MPG", 
     xlab="Weight in 1000s", ylab="MPG", pch=19)

abline(cars.regression, col="orange")
```

