---
title: "Week 5 - Test Statistics & P-values"
author: "Lacey Hartigan"
date: "2/10/21"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r  include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This week we will not do much with R other than calculate test statistics and compute p-values.

Given a one sample z-test, two-tailed, alpha level .05. 

Let's say we have a population mean of 1400 and standard dev of 80. Our sample mean is 1350 for a sample of students who took an SAT prep course. Does our sample of 28 students statistically significantly differ from the population?

First, let's write out our Null and Alternative hypotheses (in words AND in symbols):

\[
\begin{eqnarray}
H_0: \mu_{satprep} = \mu_0 \\
\end{eqnarray}
\]
There will be no difference in the means between students who took the SAT prep course and the general population of test takers.
\[
\begin{eqnarray}
H_1: \mu_{satprep} \neq \mu_0 \\ 
\end{eqnarray}
\]
There will be a difference in the means between students who took the SAT prep course and the general population of test takers.

### METHOD 1: Critical Value
Use the critical value method to determine whether the difference between students who took the SAT prep course and the general population is statistically significant.

1a. Calculate the test statistic (Z)
```{r}
(1350-1400)/(145/sqrt(28))
```
1b. Compare the test statistic to the critical value (CV) for alpha=.05, 2-tailed test (1.96). Is our test statistic MORE extreme than +/- 1.96? 

A: No. -1.82 is LESS extreme than the CV of -1.96. So, we fail to reject the null. We did NOT find a statistically significant difference.

### METHOD 2: P-Value
Let's now use the p-value method to determine whether the difference between students who took the SAT prep course and the general population is statistically significant.

2a. Calculate the p-value using pnorm. Make sure you use the correct lower.tail argument (lower.tail=T for negative test statistics; lower.tail=F for positive test statistics)
```{r}
2 * pnorm(-1.824656, lower.tail = T)
```

2b. Compare the p-value to our alpha (in this, and MOST cases, it's .05). Is p LESS than alpha? 

A: No. p is > .05. So, we fail to reject the null. We did NOT find a statistically significant difference. Note: Does this answer agree with your answer from using the CV method (IT BETTER!).

2b-1. We COULD have also entered all data into pnorm without calculating the test statistic first, right? 
```{r}
2*pnorm(1350, mean=1400, sd=145/sqrt(28), lower.tail=T)
```
Of course, this gets us the exact same answer.


### USING R FOR THE GROUP PROBLEM IN CLASS (after breakout groups)






